---
title: "iteration_and_listcols"
author: "Hanrui Li"
date: "2024-10-29"
output: github_document
---

```{r message=FALSE,warning=FALSE}
library(tidyverse)
library(rvest)

set.seed(1)
```

## Lists

You can put anything in a list.

```{r}
vec_numeric = 5:8
vec_char = c("My", "name", "is", "Jeff")
vec_logical = c(TRUE, TRUE, TRUE, FALSE)
```

```{r}
l = list(
  vec_numeric = 5:8,
  mat = matrix(1:8, nrow = 2, ncol = 4),
  vec_logical = c(TRUE, TRUE, FALSE, TRUE, FALSE, FALSE),
  summary = summary(rnorm(1000)))
```

```{r}
l

l$vec_numeric
l[[1]]
l[[1]][1:3]
mean(l[["vec_numeric"]])
```


## `for` loops

Create a new list.

```{r}
list_norms = 
  list(
    a = rnorm(20, 3, 1),
    b = rnorm(20, 0, 5),
    c = rnorm(20, 10, .2),
    d = rnorm(20, -3, 1)
  )

list_norms

is.list(list_norms)
```

```{r}
mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) == 1) {
    stop("Cannot be computed for length 1 vectors")
  }
  
  mean_x = mean(x)
  sd_x = sd(x)

  tibble(
    mean = mean_x, 
    sd = sd_x
  )
}
```

I can apply the function to each list element.

```{r}
mean_and_sd(list_norms[[1]])
mean_and_sd(list_norms[[2]])
mean_and_sd(list_norms[[3]])
mean_and_sd(list_norms[[4]])
```

Let's use a `for` loop.

```{r}
output = vector("list", length = 4)

for (i in 1:4) {
  output[[i]] = mean_and_sd(list_norms[[i]])
}
```


## `map`

```{r}
output = map(list_norms, mean_and_sd)
output
```

```{r}
output = vector("list", length = 4)

for (i in 1:4) {
  output[[i]] = median(list_norms[[i]])
}

output = map(list_norms, median)
output
```


## `map` variants

```{r}
output = map_dbl(list_norms, median, .id = "input")
# median outputs a single numeric value each time
# result is a vector instead of a list
# .id keeps the names of the elements in the input list
output
```

Produce a single data frame.

```{r}
output = map_dfr(list_norms, mean_and_sd, .id = "input")
output
```

When the function has 2 arguments

```{r}
# output = map2(input_1, input_2, \(x, y) func(arg_1 = x, arg_2 = y))

x = list(1, 1, 1)
y = list(10, 20, 30)

map2(x, y, \(x, y) x + y)
```


## List columns and operations

```{r}
listcol_df = tibble(
  name = c("a", "b", "c", "d"),
  samp = list_norms
)

# name column is a character column (character vector from data frame)
# samp column is a list column

listcol_df |>
  pull(name)

listcol_df |>
  pull(samp)

listcol_df$samp[[1]]
# first list entry
```

Apply `mean_and_sd` to the first element of the list column

```{r}
mean_and_sd(listcol_df$samp[[1]])

map(listcol_df$samp, mean_and_sd)
# apply mean_and_sd to each element
```

Use `mutate` to define a new variable in a data frame

```{r}
listcol_df = 
  listcol_df |>
  mutate(summary = map(samp, mean_and_sd))

listcol_df
```


## Revisiting NSDUH

```{r}
nsduh_table <- function(html, table_num) {
  
  table = 
    html |> 
    html_table() |> 
    nth(table_num) |>
    slice(-1) |> 
    select(-contains("P Value")) |>
    pivot_longer(
      -State,
      names_to = "age_year", 
      values_to = "percent") |>
    separate(age_year, into = c("age", "year"), sep = "\\(") |>
    mutate(
      year = str_replace(year, "\\)", ""),
      percent = str_replace(percent, "[a-c]$", ""),
      percent = as.numeric(percent)) |>
    filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
  
  table
}
```

import three tables

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

output = vector("list", 3)

for (i in c(1, 4, 5)) {
  output[[i]] = nsduh_table(nsduh_html, i)
}

nsduh_results = bind_rows(output)
```

import these data using `map`

```{r}
nsduh_results = 
  map(c(1, 4, 5), nsduh_table, html = nsduh_html) |> 
  bind_rows()
```

import using data frames and list columns

```{r}
nsduh_results= 
  tibble(
    name = c("marj", "cocaine", "heroine"),
    number = c(1, 4, 5)) |> 
  mutate(table = map(number, \(num) nsduh_table(html = nsduh_html, num))) |> 
  unnest(cols = "table")
```


## Operations on nested data

```{r message=FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

```{r}
weather_nest = 
  nest(weather_df, data = date:tmin)

weather_nest

weather_nest |> pull(name)

weather_nest |> pull(data)

unnest(weather_nest, cols = data)
```

fit the simple linear regression relating `tmax` to `tmin`

```{r}
weather_lm = function(df) {
  lm(tmax ~ tmin, data = df)
}

weather_lm(weather_nest$data[[1]])

map(weather_nest$data, weather_lm)

map(weather_nest$data, \(df) lm(tmax ~ tmin, data = df))
```

mutate to fit the model

```{r}
weather_nest = 
  weather_nest |> 
  mutate(models = map(data, weather_lm))

weather_nest
```


## Using iteration for data import

```{r message=FALSE,warning=FALSE}
full_df = 
  tibble(
    files = list.files("data/"),
    path = str_c("data/", files)
  ) |>
  mutate(data = map(path, read_csv)) |>
  unnest()
```

```{r}
tidy_df = 
  full_df |>
  mutate(
    files = str_replace(files, ".csv", ""),
    group = str_sub(files, 1, 3)) |>
  pivot_longer(
    week_1:week_8,
    names_to = "week",
    values_to = "outcome",
    names_prefix = "week_") |>
  mutate(week = as.numeric(week)) |> 
  select(group, subj = files, week, outcome)
```

```{r}
tidy_df |> 
  ggplot(aes(x = week, y = outcome, group = subj, color = group)) + 
  geom_point() + 
  geom_path() + 
  facet_grid(~group)
```


