---
title: "writing_functions"
author: "Hanrui Li"
date: "2024-10-31"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)

set.seed(1)
```

## Do something simple

```{r}
x_vec = rnorm(25, mean = 5, sd = 3)

(x_vec - mean(x_vec)) / sd(x_vec)
```

Want a function to compute z-zcores

```{r}
z_scores = function(x) {
  z = (x - mean(x)) / sd(x)
  return(z)
}

z_scores(x_vec)
```

Try my functions on some other things. These should give errors.

```{r}
z_scores(3)

# z_scores("my name is jeff") # error: non-numeric

# z_scores(iris) # error: cannot be coerced to type 'double'

z_scores(c(TRUE,TRUE,FALSE,TRUE))

z_scores(sample(c(TRUE, FALSE), 25, replace = TRUE))
```

```{r}
z_scores = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) == 1) {
    stop("Z scores cannot be computed for length 1 vectors")
  }
  
  z = mean(x) / sd(x)
  
  z
}
```


## Multiple outputs

```{r}
mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) == 1) {
    stop("Cannot be computed for length 1 vectors")
  }
  
  mean_x = mean(x)
  sd_x = sd(x)

  list(mean = mean_x, 
       sd = sd_x)
}
```

Alternatively, we might store values in a data frame.

```{r}
mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) == 1) {
    stop("Cannot be computed for length 1 vectors")
  }
  
  mean_x = mean(x)
  sd_x = sd(x)

  tibble(
    mean = mean_x, 
    sd = sd_x
  )
}
```

Check that the function works.

```{r}
x_vec = rnorm(100, mean = 3, sd = 4)
mean_and_sd(x_vec)
```


## Multiple inputs

I'd like to do this with a function.

```{r}
sim_data = tibble(
  x = rnorm(n = 100, mean = 4, sd = 3))

sim_data |> 
  summarize(
    mean = mean(x),
    sd = sd(x)
  )
```

```{r}
sim_mean_sd = function(sample_size, mu, sigma) {
  
  sim_data = tibble(
    x = rnorm(n = sample_size, mean = mu, sd = sigma),
  )
  
  sim_data |> 
    summarize(
      mean = mean(x),
      sd = sd(x)
    )
}

sim_mean_sd(100, 6, 3)
```

```{r}
sim_mean_sd = function(sample_size, mu = 3, sigma = 3) {
  
  sim_data = tibble(
    x = rnorm(n = sample_size, mean = mu, sd = sigma),
  )
  
  sim_data |> 
    summarize(
      mean = mean(x),
      sd = sd(x)
    )
}

sim_mean_sd(100, 6, 3)
```


## Learning Assignment: Loading LoTR data

```{r}
fellowship_ring = readxl::read_excel("LotR_Words.xlsx", range = "B3:D6") |>
  mutate(movie = "fellowship_ring")

two_towers = readxl::read_excel("LotR_Words.xlsx", range = "F3:H6") |>
  mutate(movie = "two_towers")

return_king = readxl::read_excel("LotR_Words.xlsx", range = "J3:L6") |>
  mutate(movie = "return_king")

lotr_tidy = bind_rows(fellowship_ring, two_towers, return_king) |>
  janitor::clean_names() |>
  pivot_longer(
    female:male,
    names_to = "sex",
    values_to = "words") |> 
  mutate(race = str_to_lower(race)) |> 
  select(movie, everything())
```

Write a function that can be used to abstract the data loading and cleaning process.

Use this function to recreate the tidied LoTR dataset.

```{r}
lotr_load_and_tidy = function(path, range, movie_name) {
  
  df = 
    readxl::read_excel(path, range = range) |>
    janitor::clean_names() |>
    pivot_longer(
      female:male,
      names_to = "sex",
      values_to = "words") |>
    mutate(
      race = str_to_lower(race),
      movie = movie_name) |> 
    select(movie, everything())
  
  df
  
}

lotr_tidy = 
  bind_rows(
    lotr_load_and_tidy("LotR_Words.xlsx", "B3:D6", "fellowship_ring"),
    lotr_load_and_tidy("LotR_Words.xlsx", "F3:H6", "two_towers"),
    lotr_load_and_tidy("LotR_Words.xlsx", "J3:L6", "return_king"))
```


## NSDUH

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

data_marj = 
  nsduh_html |> 
  html_table() |> 
  nth(1) |>
  slice(-1) |> 
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
```

Write a function to scrape review information for other tables on this page.

```{r}
nsduh_table <- function(html, table_num, table_name) {
  
  table = 
    html |> 
    html_table() |> 
    nth(table_num) |>
    slice(-1) |> 
    select(-contains("P Value")) |>
    pivot_longer(
      -State,
      names_to = "age_year", 
      values_to = "percent") |>
    separate(age_year, into = c("age", "year"), sep = "\\(") |>
    mutate(
      year = str_replace(year, "\\)", ""),
      percent = str_replace(percent, "[a-c]$", ""),
      percent = as.numeric(percent),
      name = table_name) |>
    filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
  
  table
  
}
```

Use this to get a few different tables and combine the results.

```{r}
nsduh_results = 
  bind_rows(
    nsduh_table(nsduh_html, 1, "marj_one_year"),
    nsduh_table(nsduh_html, 4, "cocaine_one_year"),
    nsduh_table(nsduh_html, 5, "heroin_one_year")
  )

nsduh_results
```








